Model Finding for Recursive Functions in SMT

### 2015-04-29 tex/main.tex
more
Most of the evaluation section.

\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
\newcommand{\cvciii}{\textsc{cvc}{\small 3}\xspace}
\newcommand{\ziii}{\textsc{z}{\small 3}\xspace}
\newcommand{\teq}{\approx}
\newcommand{\cc}[1]{#1^*}
\newcommand{\terms}{\mathbf{T}}
\newcommand{\functerms}{\mathbf{DFT}}
\newcommand{\M}{\mathcal{M}}
\def\AIF{\qtab\keyword{if}\ }
\def\THEN{\ \keyword{then}\ }
\def\AELSE{\untab\qtab\keyword{else}\ }
\def\FI{\untab}
\def\RETURN{\keyword{return}\ }
\def\ENDPROC{\untab}
\newcommand{\conv}{\mathcal{C}}

@@

@@

Say that $\Sigma$-formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$,
and the definitions in $\sfundefs{\Sigma}$ are logically consistent \ajr{formalize}.
In this section, we present a method which constructs an extended signature $\Sigma'$,
and translates $\varphi$ to equisatisfiable $\Sigma'$-formula $\psi$ on which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.

@@

The intuition behind this translation is use an uninterpreted type $\fargsort{f}$ to abstract the set of tuples that each recursive function $f$ is invoked on,
and restrict the quantification of the function definition for $f$ to a variable of type $\fargsort{f}$.
In more detail, we construct an extended signature of $\Sigma$, which we will write as $\extendsig{\Sigma}$, as follows.
For each $f \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$, 
the signature $\extendsig{\Sigma}$ includes:
\begin{enumerate}
\item[-] an uninterpreted type $\fargsort{f}$, and
\item[-] a vector of $n$ uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ of type $(\fargsort{f} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{f} \rightarrow \tau_n)$.
\end{enumerate}
%The interpretation of uninterpreted sort $\fargsort{f}$ will denote the elements (tuples) on which the function $f$ is applied.
%The role of the uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ will be discussed more in the following.
We then run the procedure $\conv$ in Figure~\ref{fig:encoding} on $\varphi$.
This calls the subprocedure $\conv_0$ on $\varphi$ and the \emph{polarity} $\ptrue$.
The procedure $\conv_0$ takes two arguments: the term $t$ to translate, and a polarity $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas, which we will call its \emph{domain constaints}.
The role of $D$ is to ensure that the (restricted) function definition for $f$ includes certain tuples in its domain, which we explain more in the following.

In the case that $t$ is an application of (interpreted) predicate $\boolop$ (e.g. a boolean connective, or equality),
we recursively call $\conv$ on its subchildren $t_i$ and polarity $\pol( \boolop, i, p )$, where $\pol$ is defined as\footnote{The negation $\neg p$ of polarity $p$ is 
$\pfalse$ if $p$ is $\ptrue$,
$\ptrue$ if $p$ is $\pfalse$, and
$\pnone$ if $p$ is $\pnone$.}:
\begin{equation*}
\pol( \boolop, i, p ) = \begin{cases}
                         \pnone & \boolop \equiv \teq \text{ or } ( \boolop \equiv \lite \text{ and } i=0 ) \\
                         \neg p & \boolop \equiv \neg \\
                         p & \text{otherwise}
                         \end{cases}
\end{equation*}
If $t$ has positive polarity,
we construct a conjunction of the result of the recursive calls with domain constraints $D_1, \ldots, D_n$.
Similarly, if $t$ has negative polarity,
we construct a disjunction of the result of the recursive calls with the negation of domain constraints $D_1, \ldots, D_n$.
It $t$ has no polarity,
then we return the union of the domain constraints $D_1 \cup \ldots \cup D_n$.

In the case that $t$ is a quantified formula $\forallf{f} \vec x. t_1$, it recursively calls $\conv$ on its body with the same polarity.
In the case that $t$ is not a function definition, 
the resulting quantified formula is reconstructed with the term $r_1$ returned by the recursive call,
and a quantified prefix is appended to the domain constraints $D_1$ returned by the recursive call.
In the case that $t$ is a function definition,
we instead construct a quantified formula over a single variable $y$ of type $\fargsort{f}$,
and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{f}( y )$.
In this case, since function definitions are a top-level conjunct, by case analysis on the return values of $\conv$, we know that $D_1$ is empty.

@@

\begin{example}
Consider the (combined) signature $\Sigma = \Sigmalia \cup \Sigma_u$,
where $\ssorts{\Sigma_u} = \{ \sortint \}$ and 
$\sfuns{\Sigma_u} = \{ s\typ{ \sortint \rightarrow \sortint }, c\typ{ \sortint } \}$.
Consider the $\Sigma$-formula $\varphi$:
\begin{equation*}
\forall_{s} x\typ{\sortint}. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
\end{equation*}
Here, $s$ is a function that returns the sum of positive numbers between $0$ and its argument $x$.
The formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma_u}$,
and states that the sum of all positive numbers between $0$ and $c$ is greater than $100$.
The signature $\extendsig{\Sigma}$ includes the sort $\fargsort{s}$,
and the uninterpreted function $\farg{s}$ of sort $\fargsort{s} \rightarrow \sortint$.
The result of $\conv( \varphi, \ptrue )$ is the pair $( \psi, \emptyset )$, where after simplification, $\psi$ is the $\extendsig{\Sigma}$-formula:
\begin{equation*}
\begin{split}
\forall y\typ{\fargsort{s}}. ite( & \farg{s}( y ) \leq 0,  \\
 & s(\farg{s}( y )) \teq 0, \\
 & s(\farg{s}( y )) \teq \farg{s}( y ) + s( \farg{s}( y )-1 ) ) \\
 & \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq \farg{s}( y )-1 )
\end{split}  
\wedge ( s( c ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq c )
\end{equation*}
The universal quantification in the resulting formula (the left disjunct) is over an uninterpreted sort $\fargsort{s}$, 
thus enabling finite model finding techniques~\cite{ReyEtAl-1-RR-13} to find a model for $\psi$ if a finite interpretation exists for $\fargsort{s}$.
Moreover, all occurrences of $y$ in this formula are beneath applications of the uninterpreted function $\farg{s}$,
implying that the formula is in the essentially uninterpreted fragment of~\cite{GeDeM-CAV-09}, for which there exists a complete instantiation procedure.
$\square$
\end{example}

@@

For benchmarks, we gathered a set of 245 benchmarks from two sources, which we will refer to as {\bf isa} and {\bf leon}.
The first ({\bf isa}) consists of the 79 benchmarks from the Isaplanner benchmark suite~\cite{DBLP:conf/itp/JohanssonDB10} that did not contain higher-order functions.
The benchmarks have been used recently as challenge problems for a variety of inductive theorem provers.
These benchmarks heavily involve recursive functions, and are limited to the combined theory of uninterpreted functions and inductive datatypes.
The second ({\bf leon}) consists of 166 benchmarks from the Leon repository~\footnote{Available at \url{https://github.com/epfl-lara/leon}.},
which were constructed from verification conditions from simple Scala programs.
These benchmarks also heavily involve recursive functions, 
and cover a wide variety of theories, including bitvectors, arrays, linear and non-linear arithmetic, and inductive datatypes.
Each of the 245 benchmarks are of the form described in Formula (~\ref{eq:wdf}) from Section~\ref{sec:prelim},
namely they each are a conjoined list of function definitions followed by a (negated) conjecture.
A majority of these benchmarks are unsatisfiable, although a handful of the benchmarks in the {\bf leon} set are satisfiable.

For each of these 245 benchmarks, we considered (up to) three randomly selected mutated forms of its conjecture $\psi$.
In particular, we considered unique conjectures that are obtained as a result of swapping a subterm of $\psi$ at one position
with another of the same type at another position.
In total, we considered 213 mutated forms of conjectures from {\bf isa}, and 427 mutated forms of conjectures from {\bf leon}.
We will call these sets {\bf isa+m} and {\bf leon+m} respectively.

In total, our benchmark set consisted of 885 benchmarks (640 mutants and 245 originals).
We considered these 885 benchmarks both before and after the translation method $\conv$ as described in Section~\ref{sec:encoding}.
We will denote the latter encoding using the suffix {\bf w}, for instance {\bf isa+w} or {\bf isa+mw}.

For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008}, 
which runs both heuristic methods for quantifier instantiation,
as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as {\bf cvc4}, {\bf cvc4+f} and {\bf cvc4+i}.
The default configuration {\bf cvc4} runs heuristic quantifier instantiation, but does not include techniques for finite model finding.
The configuration {\bf cvc4+f} runs a finite model finding procedure~\cite{ReyEtAl-1-RR-13}.
The last configuration {\bf cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.

The results are shown in Figure~\ref{fig:results}.

@@

  * future work
    * (co)inductive predicates

>>>>

\newcommand{\cvc}{\textsc{cvc}{\small 4}\xspace}
\newcommand{\cvciii}{\textsc{cvc}{\small 3}\xspace}
\newcommand{\ziii}{\textsc{z}{\small 3}\xspace}
\newcommand{\teq}{\approx}
\newcommand{\cc}[1]{#1^*}
\newcommand{\terms}{\mathbf{T}}
\newcommand{\functerms}{\mathbf{DFT}}
\newcommand{\M}{\mathcal{M}}
\def\AIF{\qtab\keyword{if}\ }
\def\THEN{\ \keyword{then}\ }
\def\AELSE{\untab\qtab\keyword{else}\ }
\def\FI{\untab}
\def\RETURN{\keyword{return}\ }
\def\ENDPROC{\untab}
\newcommand{\conv}{\mathcal{W}}

@@

\ajr{Demand for recursive functions in SMT (new draft includes define-funs-rec).  Usually not interested in showing logical consistency, there exist other tools that check e.g. termination. 
Thus, want SMT solver to check satisfiability the assumption that recursive functions are consistent (where this needs to be defined precisely).
}

\paragraph{Contributions}

\ajr{(1) Encoding, (2) Evaluation, (3) Front-end support for define-funs-rec in \cvc}

@@

Say that $\Sigma$-formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma}$,
and the definitions in $\sfundefs{\Sigma}$ are logically consistent \ajr{formalize}.
In this section, we present a method which constructs an extended signature $\Sigma'$,
and translates $\varphi$ to equisatisfiable $\Sigma'$-formula $\psi$ on which known model-finding procedures~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are applicable.
The intuition behind this translation is use an uninterpreted type to abstract the set of tuples that each recursive function $f$ is invoked on,
and restrict the quantification of the function definition for $f$ to a variable of this type.

@@

In more detail, 
say our signature $\Sigma$ contains a set of defined function symbols $\sfundefs{\Sigma} \subseteq \sfuns{\Sigma}$.
We extend the signature of $\Sigma$ to a signature $\extendsig{\Sigma}$ containing the following additional definitions.
For each $f \in \sfundefs{\Sigma}$ of type $\tau_1 \times \ldots \tau_n \rightarrow \tau$, 
the signature $\extendsig{\Sigma}$ includes:
\begin{enumerate}
\item[-] an uninterpreted type $\fargsort{f}$, and
\item[-] a vector of $n$ uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ of type $(\fargsort{f} \rightarrow \tau_1)$, $\ldots$, $(\fargsort{f} \rightarrow \tau_n)$.
\end{enumerate}
%The interpretation of uninterpreted sort $\fargsort{f}$ will denote the elements (tuples) on which the function $f$ is applied.
%The role of the uninterpreted functions $\farg{f}^1$, $\ldots$, $\farg{f}^n$ will be discussed more in the following.

Given the extended signature, we run the procedure $\conv$ in Figure~\ref{fig:encoding} on $\varphi$.
This calls the subprocedure $\conv_0$, which takes two arguments: the term $t$ to translate, and a \emph{polarity} $p$ (either $\ptrue$, $\pfalse$, or $\pnone$).
It returns a pair of the form $( r, D )$, where $r$ is a term of type $\tau$, and $D$ is a set of formulas.
%The role of $D$ is to ensure that the (restricted) function definition for $f$ includes certain tuples in its domain, which we explain more in the following.
At a high level, the translation modifies $\varphi$ in two ways.
First, it restricts the quantification on function definitions for $f$ to the corresponding uninterpreted sort $\fargsort{f}$.
Second, it augments $\varphi$ with additional constraints of the form $\exists z\typ{\fargsort{f}}. ( \vecfarg{f}( z ) \teq \vec t )$
which ensure that this restriction for the defintion $f$ covers all relevant tuples of terms,
namely the ones for which an $f$-application exists and is relevant to the satisfiability of the current branch of $\varphi$.

In the case that $t$ is an application of (interpreted) predicate $\boolop$ (e.g. a boolean connective, or equality),
we recursively call $\conv$ on its subchildren $t_i$ and polarity $\pol( \boolop, i, p )$, where $\pol$ is defined as\footnote{The negation $\neg p$ of polarity $p$ is 
$\pfalse$ if $p$ is $\ptrue$,
$\ptrue$ if $p$ is $\pfalse$, and
$\pnone$ if $p$ is $\pnone$.}:
\begin{equation*}
\pol( \boolop, i, p ) = \begin{cases}
                         \pnone & \boolop \equiv \teq \text{ or } ( \boolop \equiv \lite \text{ and } i=0 ) \\
                         \neg p & \boolop \equiv \neg \\
                         p & \text{otherwise}
                         \end{cases}
\end{equation*}
If $t$ has positive polarity,
we construct a conjunction of the result of the recursive calls with formulas $D_1, \ldots, D_n$.
Similarly, if $t$ has negative polarity,
we construct a disjunction of the result of the recursive calls with the negation of formulas $D_1, \ldots, D_n$.
It $t$ has no polarity,
then we return the union of $D_1 \cup \ldots \cup D_n$.

In the case that $t$ is a quantified formula $\forallf{f} \vec x. t_1$, it recursively calls $\conv$ on its body with the same polarity.
In the case that $t$ is not a function definition, 
the resulting quantified formula is reconstructed with the term $r_1$ returned by the recursive call,
and a quantified prefix is appended to the formula $D_1$ returned by the recursive call.
In the case that $t$ is a function definition,
we instead construct a quantified formula over a single variable $y$ of type $\fargsort{f}$,
and replace all occurrences of $\vec x$ in $r_1$ with $\vecfarg{f}( y )$.
In this case, since function definitions are a top-level conjunct, by case analysis on the return values of $\conv$, we know that $D_1$ is empty.

@@

\begin{example}
Consider the (combined) signature $\Sigma = \Sigmalia \cup \Sigma_u$,
where $\ssorts{\Sigma_u} = \{ \sortint \}$ and 
$\sfuns{\Sigma_u} = \{ s\typ{ \sortint \rightarrow \sortint }, c\typ{ \sortint } \}$.
Consider the $\Sigma$-formula $\varphi$:
\begin{equation*}
\forall_{s} x\typ{\sortint}. ite( x \leq 0, s(x) \teq 0, s( x ) \teq x + s( x-1 ) ) \wedge s( c ) > 100
\end{equation*}
Here, $s$ is a function that returns the sum of positive numbers between $0$ and its argument $x$.
The formula $\varphi$ is well-defined with respect to functions $\sfundefs{\Sigma_u}$,
and states that the sum of all positive numbers between $0$ and $c$ is greater than $100$.
This formula is satisfiable with a model that interprets $c$ as an integer greater than or equal to $14$.
Due to the universal quantification in the left disjunct,
state-of-the-art SMT techniques~\cite{GeDeM-CAV-09, ReyEtAl-1-RR-13} are unable to find a model for this formula.
The signature $\extendsig{\Sigma}$ includes the sort $\fargsort{s}$,
and the uninterpreted function $\farg{s}$ of sort $\fargsort{s} \rightarrow \sortint$.
The result of $\conv( \varphi, \ptrue )$ is the pair $( \psi, \emptyset )$, where after simplification, $\psi$ is the $\extendsig{\Sigma}$-formula:
\begin{equation*}
\begin{split}
\forall y\typ{\fargsort{s}}. ite( & \farg{s}( y ) \leq 0,  \\
 & s(\farg{s}( y )) \teq 0, \\
 & s(\farg{s}( y )) \teq \farg{s}( y ) + s( \farg{s}( y )-1 ) ) \\
 & \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq \farg{s}( y )-1 )
\end{split}  
\wedge ( s( c ) > 100 \wedge \exists z\typ{\fargsort{s}}. \farg{s}( z ) \teq c )
\end{equation*}
The universal quantification in the resulting formula is over an uninterpreted sort $\fargsort{s}$, 
thus enabling finite model finding techniques~\cite{ReyEtAl-1-RR-13} to search for a finite interpretation for $\fargsort{s}$.
Moreover, all occurrences of $y$ in this formula are beneath applications of the uninterpreted function $\farg{s}$,
implying that the formula is in the essentially uninterpreted fragment of~\cite{GeDeM-CAV-09}, for which there exists a complete instantiation procedure.
$\square$
\end{example}

@@

For benchmarks, we gathered a set of 245 benchmarks from two sources, which we will refer to as {\bf isa} and {\bf leon}.
The first ({\bf isa}) consists of the 79 benchmarks from the Isaplanner benchmark suite~\cite{DBLP:conf/itp/JohanssonDB10} that did not contain higher-order functions.
The benchmarks have been used recently as challenge problems for a variety of inductive theorem provers.
These benchmarks heavily involve recursive functions, and are limited to the combined theory of uninterpreted functions and inductive datatypes.
The second ({\bf leon}) consists of 166 benchmarks from the Leon repository~\footnote{Available at \url{https://github.com/epfl-lara/leon}.},
which were constructed from verification conditions from simple Scala programs.
These benchmarks also heavily involve recursive functions over inductive datatypes, 
and also cover a wide variety of theories, including bitvectors, arrays, and linear and non-linear arithmetic.
Each of the 245 benchmarks are of the form described in Formula (~\ref{eq:wdf}) from Section~\ref{sec:prelim},
namely they each are a conjoined list of function definitions followed by a (negated) conjecture.
A majority of these benchmarks are unsatisfiable, although a handful of the benchmarks in the {\bf leon} set are satisfiable.

For each of these 245 benchmarks, we considered (up to) three randomly selected mutated forms of its conjecture $\psi$.
In particular, we considered unique conjectures that are obtained as a result of swapping a subterm of $\psi$ at one position
with another of the same type at another position.
Note that benchmarks created in this way have a high likelihood of having small and easy-to-find counterexamples.
In total, we considered 213 mutated forms of conjectures from {\bf isa}, and 427 mutated forms of conjectures from {\bf leon}.
We will call these sets {\bf isa+m} and {\bf leon+m} respectively.

In total, our benchmark set consisted of 885 benchmarks (640 mutants plus the original 245).
We considered these 885 benchmarks both before and after the translation $\conv$ as described in Section~\ref{sec:encoding}.
We will denote the latter encoding using the suffix {\bf w}.
For instance {\bf isa+w}, contains 79 benchmarks that are the results of applying $\conv$ to each of the 79 benchmarks in {\bf isa}.

For solvers, we considered the SMT solver \ziii~\cite{de-moura-bjoerner-2008}, 
which runs both heuristic methods for quantifier instantiation~\cite{DBLP:conf/cade/MouraB07},
as well as methods for finding models for quantified formulas~\cite{GeDeM-CAV-09}.
We also considered three configurations of \cvc~\cite{barrett-et-al-2011} which we will refer to as {\bf cvc4}, {\bf cvc4+f} and {\bf cvc4+i}.
The default configuration {\bf cvc4} runs heuristic quantifier instantiation, 
but does not include techniques for finding models.
The configuration {\bf cvc4+f} runs a finite model finding procedure~\cite{ReyEtAl-1-RR-13, reynolds-et-al-2013}.
The last configuration {\bf cvc4+i} incorporates techniques for automating inductive reasoning in SMT~\cite{reynolds-kuncak-2015}.
%All configurations of \cvc incorporate techniques for conflict-based instantiation~\cite{} ..?

\begin{figure}
\centering
{
\begin{tabular}{|l|cc|cc|cc|cc|l|cc|cc|cc|cc|}
\hline                                                                
  & \multicolumn{2}{c|}{{\bf z3}}     & \multicolumn{2}{c|}{{\bf cvc4}}     & \multicolumn{2}{c|}{{\bf cvc4+f}}     & \multicolumn{2}{c|}{{\bf cvc4+i}}     & & \multicolumn{2}{c|}{{\bf z3}}     & \multicolumn{2}{c|}{{\bf cvc4}}     & \multicolumn{2}{c|}{{\bf cvc4+f}}     & \multicolumn{2}{c|}{{\bf cvc4+i}}     
\\                                                                    
  & u & s & u & s & u & s & u & s & & u & s & u & s & u & s & u & s 
\\                                                                    
\hline                                                                    
{\bf isa} & 14  & 0 & 15  & 0 & 15  & 0 & {\bf 61}  & 0 & {\bf isa+w} & 15  & 0 & 15  & 0 & 15  & 0 & 14  & 0
\\                                                                  
{\bf leon}  & 73  & 0 & 80  & 0 & 80  & 0 & {\bf 96}  & 0 & {\bf leon+w}  & 78  & 2 & 80  & 0 & 76  & {\bf 9} & 78  & 0
\\                                                                   
{\bf isa+m} & 17  & 0 & 18  & 0 & 18  & 0 & {\bf 44}  & 0 & {\bf isa+mw}  & 18  & 35  & 18  & 0 & 18  & {\bf 153} & 17  & 0
\\                                                                  
{\bf leon+m}  & 83  & 11  & 103 & 6 & 104 & 6 & {\bf 117} & 6 & {\bf leon+mw} & 98  & 75  & 98  & 6 & 95  & {\bf 169} & 98  & 6
\\                                                                  
\hline                                                             
{\bf total} & 187 & 11  & 216 & 6 & 217 & 6 & {\bf 318} & 6 & {\bf total} & 209 & 112 & 211 & 6 & 204 & {\bf 331} & 207 & 6
\\                                                                                            
\hline                                                                                            
\end{tabular}
\\
}
\caption{Number of unsat and sat responses for of all configurations on benchmarks 
before and after the translation $\conv$.}
\label{fig:results}
\end{figure}

The results are shown in Figure~\ref{fig:results}.
The results are divided into benchmarks before and after the translation $\conv$.
First, we note that in no cases did a configuration answer differently across the two encodings.
In other words, there were no cases in which \ziii or \cvc answered e.g. ``unsatisfiable" on a benchmark $\varphi$ in {\bf isa} 
and ``satisfiable" on its corresponding benchmark $\conv( \varphi )$ in {\bf isa+w}.
This provides confirming evidence that the translation $\conv$ is \emph{sound},
as we did not find a case where $\conv( \varphi )$ was unsatisfiable when $\varphi$ was satisfiable.
It also provides confirming evidence that the translation is \emph{complete} for this set of benchmarks,
as we did not find a case where $\conv( \varphi )$ was satisfiable when $\varphi$ was unsatisfiable.
This is to be expected, since the functions in our benchmark sets are intended to be logically consistent.

Examining the first encoding in isolation (the first nine columns),
the number of satisfiable responses from all configurations is quite low.
In total, {\bf z3} manages to find only 11 benchmarks to be satisfiable,
whereas each configuration of \cvc only finds 6.
This confirms the shortcomings of existing SMT techniques for finding counterexamples for benchmarks containing recursive functions.
In terms of number of unsatisfiable responses,
the configuration {\bf cvc4+i} is the clear winner, finding 318 total benchmarks to be unsatisfiable.

Now, examining the second encoding (the latter nine columns), 
we see that using the translation $\conv$ drastically improves the ability of configurations for determining satisfiability.
In total, {\bf cvc4+f} finds 331 of the 885 benchmarks to be satisfiable, including 9 benchmarks in the original {\bf leon} benchmark set.
The performance of {\bf z3} for counterexamples also improves drastically, as it solves 112 satisfiable,
including 10 that were unsolved by {\bf cvc4+f}.
Moreover, the translation $\conv$ helps \ziii for unsatisfiable responses as well,
as we see \ziii solves a total of 209 whereas it solves only 187 before the translation.
This number is 2 fewer than {\bf cvc4}, which solves 211 unsatisfiable.
Thus, in most cases the translation $\conv$ does not seriously degrade performance for unsatisfiable benchmark,
and in some cases may actually aid the solver for determining unsatisfiability
\footnote{The exception here is {\bf cvc4+i}, which makes use of type information of quantified formulas to make inductive arugments.}.

@@

  * future work
    * (co)inductive predicates?
    * quotient types??
